\begin{thebibliography}{100}
\bibitem{notes_ncm}
Chandrashekar, P. (2019). {\em Numerical methods for hyperbolic system of conservation laws}. \href{https://math.tifrbng.res.in/~praveen/pub/ncm2019.pdf}{https://math.tifrbng.res.in/~praveen/pub/ncm2019.pdf}

\bibitem{rusanov}
V.V Rusanov (1962). {\em The calculation of the interaction of non-stationary shock waves and obstacles}. USSR Computational Mathematics and Mathematical Physics, 1(2), 304-320.

\bibitem{trojak}
W. TROJAK AND F. D. WITHERDEN, {\em A new family of weighted one-parameter flux reconstruction schemes, Computers \& Fluids}, 222 (2021), p. 104918.

\bibitem{hyunh}
H. T. HUYNH, {\em A Flux Reconstruction Approach to High-Order Schemes Including Discontinuous Galerkin Methods}, Miami, FL, June 2007, AIAA.

\bibitem{vincent}
P. VINCENT, F. WITHERDEN, B. VERMEIRE, J. S. PARK, AND A. IYER, {\em Towards Green Aviation with Python at Petascale, in SC16: International Conference for High Performance Computing, Networking, Storage and Analysis}, Salt Lake City, UT, Nov. 2016, IEEE, pp. 1-11.

\bibitem{trixi}
Schlottke-Lakemper, M., Gassner, G. J., Ranocha, H., Winters, A. R., \& Chan, J. (2024). {\em Trixi.jl (v0.8.3)}. Zenodo. \href{https://doi.org/10.5281/zenodo.12683615}{https://doi.org/10.5281/zenodo.12683615}

\bibitem{rma}
Hoefler, T., Balaji, P., Gropp, W., \& Thakur, R. (2016, November). \textit{Advanced MPI Programming} [Tutorial] \href{https://web.cels.anl.gov/ thakur/sc16-mpi-tutorial/slides.pdf}{https://web.cels.anl.gov/ thakur/sc16-mpi-tutorial/slides.pdf}

\bibitem{rmabook}
Hoefler, T., Gropp, W., Thakur, R., \& Lusk, E. (2014). \textit{Using advanced MPI: Modern features of the message-Passing-Interface.} MIT Press.

\bibitem{arpit}
Babbar, A., \& Chandrashekar, P. (2024). \textit{Lax-Wendroff Flux Reconstruction on adaptive curvilinear meshes with error based time stepping for hyperbolic conservation laws.} arXiv e-prints, arXiv:2402.11926.

\end{thebibliography}